Computer-Assisted Analysis of the \\ Anderson-Hájek Ontological Controversy

### 2014-10-30 Papers/InProgress/LogicAndReligionCongress/Abstract.tex
Make cosmetic changes

For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold for \emph{Anderson's Emendation} (cf. Appendix \ref{apx:Anderson}): 
%
the theorems T1, C and T3' can be quickly automated (in logics \K, \K and \KB, respectively);
%
the axioms A4 and A5 are proven redundant\footnote{
  An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is entailed by $S$. 
} (the former in
logic \KFourB and the latter already in \K);
%
a trivial countermodel (consisting of two worlds and 
two individuals) for MC is generated by Nitpick (for all mentioned logics); all axioms and definitions are shown to be mutually consistent.

The redundancy of A4 and A5 is particularly controversial. Magari \cite{Magari1988} claimed that A4 and A5 are superfluous\footnote{
  An axiom $A$ is superfluous w.r.t. a set of axioms $S$ iff T3 is entailed by $S \setminus \{ A \}$.
}, arguing that T3 is true in all models of the other axioms and definitions by Gödel. \citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, refuted Magari's claim, but claimed that A4 and A5 are indeed superfluous under the assumption of an additional axiom (PEP, pointwise equality for positiveness), which Magari probably assumed silently. Moreover, \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant. \citet[footnote 1 in p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics). Our results show that Hájek was originally right, under both constant and varying domain semantics.

Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
  A4 and A5 are used by \citet[p.~11]{Hajek2002} (page 11) in the proofs of, respectively, Lemma 4 and Theorem 4.
} 
of A4 and A5, as well as varying domain semantics, in his new
emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
\ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
axiom H:A12. Surprisingly, the computer-assisted formalization of
\AOEH\ shows that A4 and A5 are still superfluous (not needed for proving T3). Moreover, A4 and A5 are independent\footnote{
  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
} of the other axioms and definitions. Therefore, A4 and A5 are not redundant, despite their superfluousness.

@@

Additionally, Anderson's 14th footnote \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers. Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant). Unfortunately then, a countermodel for T3 can be found.

The controversy over the superfluousness of A4 and A5 indicates a trend to reduce the ontological argument to its bare essentials. In this regard, 
already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable. These claims have been confirmed by the automated theorem provers (in logic \KFourB). Within the same trend, the alternative proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal}) achieves a high level of minimality. He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being. He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1, A2, A3 and A4. This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K. A4 can be proved in logic \KT (i.e. assuming reflexivity of the accessibility relation). For constant domain semantics, proving D1 is possible in logic \KFour, whereas for varying domain semantics, a counter model can be found even in logic \SFive. Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K. The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB. Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with certain additional presuppositions, to eliminate A5 from his theory as well. Without any additional presupposition, the automated reasoners show that A5 is actually not superfluous. All these results, with the exception of the aforementioned countermodel for D1, hold for both constant and varying domain semantics.

@@

\section{Scott's version of G\"odel's ontological argment} \label{apx:Goedel} 
\begin{itemize}
\item[A1] Either a property or its negation is positive, but not
  both:
  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
\item[D1] A \emph{God-like} being possesses all positive properties: 
  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
\item[A3]  The property of being God-like is positive: 
  $$\hol{P(G)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
\item[T2]  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
  \exq y \varphi(y)]}$$
\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
%
\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}

@@

% \begin{figure}[t]
% \noindent \framebox[\columnwidth][r]{
% \begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[A:A1] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
\item[A3']  The property of being God-like is positive: 
  $$\hol{P(G_A)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
\item[T2']  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y \varphi(y)]}$$
\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
%
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
\end{itemize}
% \end{minipage}
% } \vskip-.5em
% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
% \end{figure}

@@

% \begin{figure}[t]
% \noindent \framebox[\columnwidth][r]{
% \begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
\item[A3']  The property of being God-like is positive: 
  $$\hol{P(G_H)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y [\varphi(y)]]}$$
\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$

@@

\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
% 
\end{itemize}
% \end{minipage}
% } \vskip-.5em
% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
% \end{figure}

\section{Hájek's Second Emendation \AOEHH} \label{apx:Hajek2}
% \begin{figure}[t]
% \noindent \framebox[\columnwidth][r]{
% \begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
\item[H:A3]  The property of being God-like and existing actually is positive: 
  $$\hol{P(G_A \wedge E)}$$
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
% 
\end{itemize}
% \end{minipage}
% } \vskip-.5em
% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
% \end{figure}

\section{Hájek's Third Emendation \AOEHHH} \label{apx:Hajek3}
% \begin{figure}[t]
% \noindent \framebox[\columnwidth][r]{
% \begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
\item[D4]  A property is positive\textsuperscript{\#} iff it is necessarily implied by a positive property: 
  $$\hol{P^\#(\varphi) \equiv \exq \psi[P(\psi) \wedge \nec \allq x[\psi(x) \imp \varphi(x)]]}$$
\item[A3']  The property of being God-like is positive: 
  $$\hol{P(G_H)}$$
\item[H:A4]  Positive\textsuperscript{\#} properties are necessarily positive\textsuperscript{\#}: 
  $$\hol{\allq \varphi [P^\#(\varphi) \imp \Box \; P^\#(\varphi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y [\varphi(y)]]}$$
\item[H:A5] Necessary existence is a positive\textsuperscript{\#} property: $$\hol{P^\#(\NE_A)}$$ 
% 
\end{itemize}
% \end{minipage}
% } \vskip-.5em
% \caption{Anderson's Emendation \cite{ToDo:Anderson}.\label{fig:anderson}} 
% \end{figure}

%\clearpage

@@

% \begin{figure}[t]
% \noindent \framebox[\columnwidth][r]{
% \begin{minipage}{.94\columnwidth}\small
In Bjordal's emendation $G$ (God-like) is taken as primitive and $P$ (Positive) is defined (cf. definition D).
\begin{itemize}
\item[B:D1] A formulas $\phi$ is positive iff it is necessarily the case
  that anything which is God-like has the property $\phi$.
  $$\hol{P(\phi) \equiv \nec \allq x (G(x) \imp \phi(x))}$$ 
\item[B:L1] D is logically equivalent in S4 with the union of G\"odel's definition D1 and axioms A2, A3 and A4.
  $$\hol{D \biimp D1 \wedge A2 \wedge A3 \wedge A4}$$
  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
\item[B:D2] $\phi$ is a maximal composite of object $x$'s positive properties iff $x$ has $\phi$ and $\phi$ is positive and all positive properties $\psi$ which $x$ has are such that is necessarily the case that all objects which have $\phi$ also have $\psi$.
  $$\hol{MCP(\phi,x) \equiv (\phi(x) \wedge P(\phi)) \wedge \allq \psi ((\psi(x) \wedge P(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
\item[B:D3] $x$ has the $N$-property iff x is such that if $\phi$ is a maximal composite of $x$'s positive properties then it is necessary that some object $y$ has the property $\phi$.
  $$\hol{N(x) \equiv \allq \phi (MCP(\phi,x) \imp \nec \allq y \phi(y))}$$
\item[A:A1] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
\item[A5'] The $N$-property is positive.
 $$\hol{P(N)}$$
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}
% \end{minipage}
% } \vskip-.5em
% \caption{Frode's Alternative \cite{ToDo:Frode}.\label{fig:frode}} 
% \end{figure}

%\clearpage
\end{appendix}

>>>>

For both \textbf{constant domain semantics} and \textbf{varying domain semantics}, the following results hold for \emph{Anderson's Emendation} (cf. Appendix \ref{apx:Anderson}): 
%
the theorems T1, C and T3' can be quickly automated (in logics \K, \K and \KB, respectively);
%
the axioms A4 and A5 are proven redundant\footnote{
  An axiom $A$ is redundant w.r.t. a set of axioms $S$ iff $A$ is entailed by $S$. 
} (the former in
logic \KFourB and the latter already in \K);
%
a trivial countermodel (with two worlds and 
two individuals) for MC is generated by Nitpick (for all mentioned logics); all axioms and definitions are shown to be mutually consistent.

The redundancy of A4 and A5 is particularly controversial. Magari \cite{Magari1988} claimed that A4 and A5 are superfluous\footnote{
  An axiom $A$ is superfluous w.r.t. a set of axioms $S$ iff T3 is entailed by $S \setminus \{ A \}$.
}, arguing that T3 is true in all models of the other axioms and definitions by Gödel. \citet[p.~5-6]{Hajek_Magari_and_others_1996} investigated this further, and claimed that Magari's claim is not valid, but is nevertheless true  under additional silent assumptions by Magari. Moreover, \citet[p.~2]{Hajek_Magari_and_others_1996} cites his earlier work\footnote{Although \citep{Hajek_der_Mathematiker_2001} precedes \citep{Hajek_Magari_and_others_1996} in writing, it was published only 5 years later, in German.} \citep{Hajek_der_Mathematiker_2001}, where he claims (in Theorem 5.3) that for Anderson's emended theory \citep{anderson90:_some_emend_of_goedel_ontol_proof}, A4 and A5 are not only superfluous, but also redundant. \citet[footnote 1 in p.~1]{AndersonGettings}, in a footnote, rebutted Hájek's claim, arguing that the redundancy of A4 and A5 holds only under constant domain semantics, while Anderson's emended theory ought to be taken under Cocchiarella's semantics \citep{Cocchiarella} (a varying domain semantics). Our results show that Hájek was originally right, under both constant and varying domain semantics.

Nevertheless, \citet[p.~7]{Hajek2002} acknowledges Anderson's rebuttal, and apparently accepts it, as evidenced by  his use\footnote{
  A4 and A5 are used by \citet[p.~11]{Hajek2002} in, respectively, Lemma 4 and Theorem 4.
} 
of A4 and A5, as well as varying domain semantics, in his new
emendation (named \AOEH \citep[sec.~4]{Hajek2002}, cf. Appendix
\ref{apx:Hajek1}), which replaces Anderson's A:A1 and A2 by a simpler
axiom H:A12. Surprisingly, the computer-assisted formalization of
\AOEH\ shows that A4 and A5 are still superfluous. Moreover, A4 and A5 are independent\footnote{
  An axiom $A$ is \emph{independent} of a set of axioms $S$ iff there are models of $S$ where $A$ is true and other models of $S$ where $A$ if false. 
} of the other axioms and definitions. Therefore, A4 and A5 are not redundant, despite their superfluousness.

@@

Additionally, Anderson \citep[footnote 14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only the quantifiers in T3' and in A:D2 need to be interpreted as actualistic quantifiers, while others may be taken as possibilistic quantifiers. Our computer-assisted study of this mixed variant shows that A4 is still redundant, but A5 becomes independent (hence not redundant). Unfortunately, a countermodel for T3 can then be found.

The controversy over the superfluousness of A4 and A5 indicates a trend to reduce the ontological argument to its bare essentials. In this regard, 
already \citet[p.~7]{anderson90:_some_emend_of_goedel_ontol_proof} indicates that, by taking a notion of \emph{defective} as primitive and defining the notion of \emph{positive} upon it, axioms A:A1, A2 and A4 become derivable. These claims have been confirmed by the automated theorem provers (in logic \KFourB). Within the same trend, the alternative proposed by \citet{bjordal99} (cf. Appendix \ref{apx:Bjordal}) achieves a high level of minimality. He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as those properties necessarily possessed by every God-like being. He then briefly indicates (B:L1) that B:D1 is logically equivalent, under modal logic \SFour, to the conjunction of D1, A2, A3 and A4. This has been confirmed in the computer-assisted formalization: A2 and A3 can be quickly automatically derived in logic \K. A4 can be proved in logic \KT (i.e. assuming reflexivity of the accessibility relation). For constant domain semantics, proving D1 is possible in logic \KFour, whereas for varying domain semantics, a counter model can be found even in logic \SFive. Conversely, the proof that B:D1 is entailed by D1, A2, A3 and A4 is possible already in logic \K. The provers also show that theorem T3 follows from B:D1, B:D2, B:D3, A:A1 and A5 already in logic \KB. Bjordal's last paragraph briefly mentions Hájek's ideas about the superfluousness of A5 and claims that it is possible, with (unclear) additional modifications of the definitions, to eliminate A5 from his theory as well. Without any additional modification, the automated reasoners show that A5 is actually not superfluous. All these results, with the exception of the aforementioned countermodel for D1, hold for both constant and varying domain semantics.

@@

\section{Scott's version of G\"odel's ontological argument} \label{apx:Goedel}

\framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[A1] Either a property or its negation is positive, but not
  both:
  $$\hol{\allq \varphi [P(\neg \varphi) \biimp \neg P(\varphi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
\item[D1] A \emph{God-like} being possesses all positive properties: 
  $$\hol{G(x) \equiv \forall \varphi [P(\varphi) \imp \varphi(x)]}$$ 
\item[A3]  The property of being God-like is positive: 
  $$\hol{P(G)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
\item[D2] An \emph{essence} of an individual is a property possessed by it and necessarily implying any of its properties: $$\hol{\ess{\varphi}{x} \equiv \varphi(x) \wedge \allq
  \psi (\psi(x) \imp \nec \allq y (\varphi(y) \imp \psi(y)))}$$ 
\item[T2]  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G(x) \imp \ess{G}{x}]}$$
\item[D3] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE(x) \equiv \allq \varphi [\ess{\varphi}{x} \imp \nec
  \exq y \varphi(y)]}$$
\item[A5] Necessary existence is a positive property: $$\hol{P(\NE)}$$ 
\item[L1] If a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\exq x G(x) \imp \nec \exq y G(y)}$$
\item[L2] If possibly a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\pos \exq x G(x) \imp \nec \exq y G(y)} $$
%
\item[T3] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}
\end{minipage}
} %\vskip-.5em

\clearpage

@@

\framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[A:A1] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
\item[A2] A property necessarily implied by a
  positive property is positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp P(\psi)]}$$
\item[T1] Positive properties are possibly exemplified: 
  $$\hol{\allq \varphi [P(\varphi) \imp \pos \exq x \varphi(x)]}$$ 
\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
\item[A3']  The property of being God-like is positive: 
  $$\hol{P(G_A)}$$
\item[C\phantom{1}] Possibly, a God-like being exists: $$\hol{\pos \exq x G(x)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
\item[T2']  Being God-like is an essence of any
  God-like being: $$\hol{\allq x [G_A(x) \imp \essA{G_A}{x}]}$$
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y \varphi(y)]}$$
\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$ 
\item[L1'] If a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\exq x G_A(x) \imp \nec \exq y G_A(y)}$$
\item[L2'] If possibly a god-like being exists, then necessarily a god-like being exists: 
  $$\hol{\pos \exq x G_A(x) \imp \nec \exq y G_A(y)} $$
%
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$ 
\end{itemize}
\end{minipage}
} %\vskip-.5em

\clearpage

@@

\framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
\item[A3']  The property of being God-like is positive: 
  $$\hol{P(G_H)}$$
\item[A4]  Positive properties are necessarily positive: 
  $$\hol{\allq \varphi [P(\varphi) \imp \Box \; P(\varphi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y [\varphi(y)]]}$$
\item[A5'] Necessary existence is a positive property: $$\hol{P(\NE_A)}$$

@@

\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_H(x)}$$
% 
\end{itemize}
\end{minipage}
} %\vskip-.5em

\clearpage

\section{Hájek's Second Emendation \AOEHH} \label{apx:Hajek2}

\framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[A:D1] A \emph{God-like} being necessarily possesses those and only those properties that are positive: 
  $$\hol{G_A(x) \equiv \forall \varphi [P(\varphi) \biimp \nec \varphi(x)]}$$ 
\item[H:A3]  The property of being God-like and existing actually is positive: 
  $$\hol{P(G_A \wedge E)}$$
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G_A(x)}$$
% 
\end{itemize}
\end{minipage}
} %\vskip-.5em

\section{Hájek's Third Emendation \AOEHHH} \label{apx:Hajek3}

\framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
\begin{itemize}
\item[H:A12] The negation of a property necessarily implied by a
  positive property is not positive:
  $$\hol{\allq \varphi \allq \psi [(P(\varphi) \wedge \nec \allq x [\varphi(x)
  \imp \psi(x)]) \imp \neg P(\neg \psi)]}$$
\item[H:D1] A \emph{God-like} being necessarily possesses those and only those properties that are necessarily implied by a positive property: 
  $$\hol{G_H(x) \equiv \forall \varphi [\nec \varphi(x) \biimp \exq \psi[P(\psi) \wedge \nec \allq x [\psi(x)  \imp \varphi(x))]]}$$ 
\item[D4]  A property is positive\textsuperscript{\#} iff it is necessarily implied by a positive property: 
  $$\hol{P^\#(\varphi) \equiv \exq \psi[P(\psi) \wedge \nec \allq x[\psi(x) \imp \varphi(x)]]}$$
\item[A3']  The property of being God-like is positive: 
  $$\hol{P(G_H)}$$
\item[H:A4]  Positive\textsuperscript{\#} properties are necessarily positive\textsuperscript{\#}: 
  $$\hol{\allq \varphi [P^\#(\varphi) \imp \Box \; P^\#(\varphi)]}$$ 
\item[A:D2] An \emph{essence} of an individual is a property that necessarily implies those and only those properties that the individual has necessarily: $$\hol{\essA{\varphi}{x} \equiv \allq
  \psi [\nec \psi(x) \biimp \nec \allq y (\varphi(y) \imp \psi(y))]}$$ 
\item[D3'] \emph{Necessary existence} of an individual is the necessary exemplification of all its essences: 
  $$\hol{\NE_A(x) \equiv \allq \varphi [\essA{\varphi}{x} \imp \nec
  \exq y [\varphi(y)]]}$$
\item[H:A5] Necessary existence is a positive\textsuperscript{\#} property: $$\hol{P^\#(\NE_A)}$$ 
% 
\end{itemize}
\end{minipage}
} %\vskip-.5em

\clearpage

@@

\framebox[\columnwidth][r]{
\begin{minipage}{.94\columnwidth}\small
$G$ (God-like) is taken as primitive and $P$ (Positive) is defined (cf. definition D).

\medskip

\begin{itemize}
\item[B:D1] A formulas $\phi$ is positive iff it is necessarily the case
  that anything which is God-like has the property $\phi$.
  $$\hol{P(\phi) \equiv \nec \allq x (G(x) \imp \phi(x))}$$ 
\item[B:L1] D is logically equivalent in S4 with the union of G\"odel's definition D1 and axioms A2, A3 and A4.
  $$\hol{D \biimp D1 \wedge A2 \wedge A3 \wedge A4}$$
  The proof splits into the two implication directions B:L1$^\rightarrow$ and B:L1$^\leftarrow$. B:L1$^\rightarrow$ can be further split into four single steps.
\item[B:D2] $\phi$ is a maximal composite of object $x$'s positive properties iff $x$ has $\phi$ and $\phi$ is positive and all positive properties $\psi$ which $x$ has are such that is necessarily the case that all objects which have $\phi$ also have $\psi$.
  $$\hol{MCP(\phi,x) \equiv (\phi(x) \wedge P(\phi)) \wedge \allq \psi ((\psi(x) \wedge P(\psi)) \imp \nec \allq y (\phi(y) \imp \psi(y)))}$$
\item[B:D3] $x$ has the $N$-property iff x is such that if $\phi$ is a maximal composite of $x$'s positive properties then it is necessary that some object $y$ has the property $\phi$.
  $$\hol{N(x) \equiv \allq \phi (MCP(\phi,x) \imp \nec \allq y \phi(y))}$$
\item[A:A1] If a property is positive, its negation is not positive:
  $$\hol{\allq \varphi [P(\varphi) \imp \neg P(\neg \varphi)]}$$ 
\item[A5'] The $N$-property is positive.
 $$\hol{P(N)}$$
\item[T3'] Necessarily, a God-like being exists: $$\hol{\nec \exq x G(x)}$$ 
\end{itemize}
\end{minipage}
} %\vskip-.5em

\vskip 8em

\end{appendix}
